---
title: "Advertising Analysis"
output: html_document
---

## Defining the Question

A Kenyan entrepreneur has created an online cryptography course and would want to advertise it on her blog. She currently targets audiences originating from various countries. In the past, she ran ads to advertise a related course on the same blog and collected data in the process. She would now like to employ your services as a Data Science Consultant to help her identify which individuals are most likely to click on her ads. 

### a)Specifying the Question

1. Identify which Individuals are most likely to click on the client's Ads

### b)The metric of Success
Identify which Individuals are most likely to click on the ads based on the data acquired

### c)The Context
Online marketing of a Cryptography Class by using Ads

### d)Experimental Design
1. Load Data
2. Check the Data
3. Tidy the Data
4. Analyze the Data


## Reading the Data
```{r }
# plotting libraries
library("dplyr")
library("ggplot2")

```
```{r }
# Load the dataset
advertising <- read.csv("~/Documents/R markdowns/advertising.csv")
head(advertising)

```
## Checking the Data
```{r }
# The shape of the dataset
dim(advertising)
```

```{r }
# The Data Types of 
str(advertising)
```


## Tidying the Data
```{r }
#Missing Data
colSums(is.na(advertising))
# No missing values
```

```{r }
#Duplicated Values 
advert <- unique(advertising)
dim(advert)
# No duplicated values
```


```{r }
#Outliers

boxplot(advert$Daily.Time.Spent.on.Site,xlab="Daily Time Spent on Site",main="Boxplot on Daily Time Spent on Site")
boxplot(advert$Age, xlab="Age",main="Boxplot on Age")
boxplot(advert$Area.Income,xlab="Area Income",main="Boxplot on Area Income")
boxplot(advert$Daily.Internet.Usage, xlab="Daily Internet Usage", main="Boxplot of Daily Internet Usage")
boxplot(advert$Male, xlab="Male", main="Boxplot of Male")
boxplot(advert$Clicked.on.Ad, xlab="Clicked on Ad", main="Boxplot of Clicked on Ad")

# The number of outliers
boxplot.stats(advert$Area.Income)$out
sum(table(boxplot.stats(advert$Area.Income)$out))

```
The dataset has no missing values or Duplicated values. From the boxplots we see that the Area Income is the only numeric column with outliers, there are 8 columns

Outliers were not removed to see their role in the analysis

## Exploratory Data Analysis

### Univariate
#### Numeric Columns
```{r }
#Numeric Values
num = advert[,c(1,2,3,4,7,10)]
summary(num)
```

```{r }
#Mean
mean(advert$Daily.Time.Spent.on.Site)
mean(advert$Age)
mean(advert$Area.Income)
mean(advert$Daily.Internet.Usage)
mean(advert$Male)
mean(advert$Clicked.on.Ad)
```

```{r }
# Median
median(advert$Daily.Time.Spent.on.Site)
median(advert$Age)
median(advert$Area.Income)
median(advert$Daily.Internet.Usage)
median(advert$Male)
median(advert$Clicked.on.Ad)
```

```{r }
# Mode
getMode <- function(v){
  uniqv <- unique(v)
  uniqv[which.max(tabulate(match(v, uniqv)))]
}
getMode(advert$Daily.Time.Spent.on.Site)
getMode(advert$Age)
getMode(advert$Area.Income)
getMode(advert$Daily.Internet.Usage)
getMode(advert$Male)
getMode(advert$Clicked.on.Ad)
```

```{r }
#Variance
var(advert$Daily.Time.Spent.on.Site)
var(advert$Age)
var(advert$Area.Income)
var(advert$Daily.Internet.Usage)
var(advert$Male)
var(advert$Clicked.on.Ad)

```

```{r }

# Standard Deviation
sd(advert$Daily.Time.Spent.on.Site)
sd(advert$Age)
sd(advert$Area.Income)
sd(advert$Daily.Internet.Usage)
sd(advert$Male)
sd(advert$Clicked.on.Ad)
```

```{r }
# Quantiles
quantile(advert$Daily.Time.Spent.on.Site)
quantile(advert$Age)
quantile(advert$Area.Income)
quantile(advert$Daily.Internet.Usage)
quantile(advert$Male)
quantile(advert$Clicked.on.Ad)
```

#### Histogram
```{r histogram, echo=FALSE}
hist(advert$Daily.Time.Spent.on.Site, breaks = 20,main ="Daily Time Spent on Site", col = "dodgerblue")
hist(advert$Age, breaks = 10,main = "Age",col = "dodgerblue")
hist(advert$Area.Income, breaks = 20, main = "Area Income",col = "dodgerblue")
hist(advert$Daily.Internet.Usage,breaks = 20,main = "Daily Internet Usage",col = "dodgerblue")
hist(advert$Male,breaks = 10,main = "Male",col = "dodgerblue")
hist(advert$Clicked.on.Ad,breaks = 10,main = "Clicked on Ad",col = "dodgerblue")
```
From the Male and the Clicked on Ad are categorical in nature

The distribution of Age  is skewed to the left, the Area Income and Daily Time Spent on Site distribution is skewed to the right

The Most frequent age is between approximately 25years and 40 years


#### BarPlots
```{r Barplots, echo=FALSE}
# How many Males Vs. Females
male <- table(advert$Male)
barplot(male,main = "Male",col = topo.colors(2),ylim = c(0, 800))
legend("topright",inset = .02, title="Gender",
       c("Female","Male"), fill=topo.colors(2), cex=0.8)

# No. of Clicked vs. Not Clicked
clicked <- table(advert$Clicked.on.Ad)
barplot(clicked,main = "Clicked on Ad",col = topo.colors(2), ylim = c(0,800))
legend("topright",inset = .02, title="Clicked on Ad",
       c("Not Clicked","Clicked"), fill=topo.colors(2), cex=0.8)

# the most popular countries
par(las=2, cex.axis=0.7)
country <- table(advert$Country)
barplot(sort(country[1:40], decreasing = TRUE), main = "Country",col = terrain.colors(20))

# The most popular Age
par(las=2)
age <- table(advert$Age)
barplot(sort(age[1:20], decreasing = TRUE), main = "Age",col = terrain.colors(20))

# group by gender/Male
# Gender that Spends more Time on the Internet
by_time <- advert %>% 
  group_by(Male) %>% 
  summarise(Total.Time.Spent.on.Site = sum(Daily.Time.Spent.on.Site))
p <- ggplot(by_time, aes(x = factor(Male), y = Total.Time.Spent.on.Site, fill = factor(Male)))+geom_bar(stat="identity")
p + scale_fill_discrete(name = "Male", labels = c("Female","Male"))+ labs(title="Gender that spends more time on the Internet", x="Gender")


# Group by country, Country with more clicks on Ad
by_country <- advert %>% group_by(Country) %>% summarise(clicked.ad =sum(Clicked.on.Ad[Clicked.on.Ad == 1]))

# select the first 20
rows <- by_country[1:20,]
rows

c <- ggplot(rows, aes(x = reorder(Country,clicked.ad), y=clicked.ad)) + geom_col() +coord_flip() +  geom_bar(stat="identity", fill="dodgerblue")
c + labs(title="Country with Highest Clicks on Ads", x="Countries", y="Clicked Ads")

# Females that click on ads
by_gender <- advert %>% group_by(Clicked.on.Ad) %>% summarise(gender = length(Male[Male == 0]))
by_gender

females <- ggplot(by_gender, aes(x = factor(Clicked.on.Ad), y = gender, fill=factor(Clicked.on.Ad))) + geom_bar(stat="identity")
females + scale_fill_discrete(name = "Ad Clicked", labels = c("Not Clicked","Clicked"))+ labs(title="Females that Clicked vs Not Clicked", x="Clicked on Ad", y="No. of Females")

# Males that clicked on ads

by_males <- advert %>% group_by(Clicked.on.Ad) %>% summarise(gender = length(Male[Male == 1]))
by_males

males <- ggplot(by_males, aes(x = factor(Clicked.on.Ad), y = gender, fill=factor(Clicked.on.Ad))) + geom_bar(stat="identity")
males + scale_fill_discrete(name = "Ad Clicked", labels = c("Not Clicked","Clicked"))+ labs(title="Males that Clicked vs Not Clicked", x="Clicked on Ad", y="No. of Males")

```
```{r }
# Ad topic Line
value.count <- table(advert$Ad.Topic.Line)
head(value.count)
sum(value.count)

# Only one occurrence of every topic line
```
From the Male column, 0 means Female and 1 means Male, we have more Females than Males
In the Clicked on Ad column, number of clicked versus not clicked are equal

The most popular countries are Afghanistan, Australia, Albania, Bahamas are the top 4 countries
The most popular age is 31 years
Females spend more time in the Internet compared to Males and with more Females spending time on the Internet, we see that most of them click on Ads compared to Males

But there is still a significant number of Females who don't click on Ads.

The Countries that had the highest number in clicked ads was Australia, Afghanistan and Bahamas

The Ad Topic Line are there is only one occurrence of each.

### Bivariate

#### Covariance
```{r covariance}
numeric_col <- advert[,c(1,2,3,4,7,10)]
head(numeric_col,4)

# Covariance
covariance_matrix = cov(numeric_col)
View(round(covariance_matrix,2))
```

#### Correlation
```{r correlation}
# Correlation Matrix
correlation_matrix = cor(numeric_col)
View(round(correlation_matrix,2))
```
Clicked on Ad column has strong negative correlation with Daily Time Spent on Internet and Daily Internet Usage
It has a moderate negative Correlation with Area Income and a Moderate Positive Correlation with Age

Therefore this columns
1.Daily Time Spent on Internet
2.Daily Internet Usage
3.Area Income
4.Age
Can we used to determine if a an Ad will be clicked on or not

#### Scatter Plots

```{r scatter plot, echo=FALSE}
area.income <- advert$Area.Income
internet.usage <- advert$Daily.Internet.Usage
time.spent <- advert$Daily.Time.Spent.on.Site

plot(area.income, internet.usage, xlab="Area Income",ylab = "Daily Internet Usage",main = "Area Income vs Daily Internet Usage")
plot(area.income,time.spent,xlab = "Area Income",ylab = "Daily Time Spent on the Internet",main="Area Income vs Daily Time Spent on the Internet")
plot(time.spent,internet.usage, xlab="Daily Time spent", ylab="Daily Internet Usage")
```

The more the Area Income the more time the user spent on the Internet and also the the more internet used

## Conclusion
The Client could concentrate on between the 35-40 age group and mostly Female users.

I also recommend that the Client could also run Ads that are more appealing to the Male the users so that they will click on the Ads more
